---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "80%"
)
```

# bayeslinear

<!-- badges: start -->
<!-- badges: end -->

<tt>bayeslinear</tt> provides tools to perform a Bayes Linear analysis.

## Installation

<tt>bayeslinear</tt> is currently in development; install the latest version from [GitHub](https://github.com/) with:

``` r
install.packages("devtools")
devtools::install_github("astfalckl/bayeslinear")
```

# Vanilla Bayes Linear Analysis

## Creating Belief Structures

A belief structure is the fundamental unit of information that we require to conduct a BL analysis blah blah blah. It requires the specification of $\mathbb{E}(X)$, $\mathbb{E}(D)$, $\mathrm{var}(X)$, $\mathrm{var}(D)$ and $\mathrm{cov}(X,D)$ to fully specify the inner product space. The methods in <tt>bayeslinear</tt> are based on the creation of a belief structure object through <tt>bs()</tt>. For example the one-dimensional example in Goldstein and Wooff (2007) is created as

```{r, warning = FALSE, message = FALSE}
library(bayeslinear)
library(ggplot2)
library(CVXR)
library(dplyr)
```


```{r}
E_X <- 1
E_D <- 2

var_X <- var_D <- 1
cov_XD <- 0.6

one_dimension <- bs(E_X, E_D, cov_XD, var_X, var_D)
print(one_dimension)
```

## Adjusting Belief Structures

A <tt>bs</tt> object is adjusted by some data <tt>D</tt> via the <tt>adjust</tt> method. An adjusted belief structure <tt>adj_bs</tt> is returned. 

```{r}
D <- 0
adjust(one_dimension, D)
```

## Summaries of Adjusted Belief Structures

Methods for calculating the adjustment resolutions and performing a canonical analysis on either a <tt>bs</tt> or an <tt>adj_bs</tt> object are <tt>resolution</tt> and <tt>canonical</tt>. For example, calculated on the oral glucose data (this belief structure ships with the package). Note, we don't have to observe the adjustment to calculate the resolutions and canonical directions; they are a product of our prior beliefs.

```{r}
resolution(oral_glucose)
canonical(oral_glucose)
```

# Generalised Bayes Linear Analysis

## Defining and adjusting a GBL Belief Structure

A GBL belief structure (<tt>gbl_bs</tt>) is constructed similarly to a standard belief structure object with the addition of a <tt>constraint</tt> argument that defines the solution space $\mathcal{C}$. <tt>bayeslinear</tt> assumes $\mathcal{C}$ is convex and uses <tt>CVXR</tt> to efficiently calculate the projection into $\mathcal{C}$. Consider the completely arbitrary example below.

```{r}
nx <- 2
nd <- 1

E_X <- matrix(c(0,0))
E_D <- matrix(c(0))
var_X <- matrix(c(0.6, 0.3, 0.3, 0.6), nrow = nx)
cov_XD <- matrix(c(0.4, -0.1), nrow = nx)
var_D <- matrix(c(1), nrow = nd)

gbl_bivar <- gbl_bs(
  E_X, E_D, cov_XD, var_X, var_D, 
  quote(list(Ec_adj >= 0))
)
gbl_bivar
```

Note the <tt>constraint</tt> argument is an expression being passed through to <tt>CVXR</tt> and so must be wrapped in <tt>quote()</tt>. Adjusting a <tt>gbl_bs</tt> object is the same as for vanilla Bayes linear with a few extra outputs.

```{r}
D <- c(-3)
bivar_adjust <- adjust(gbl_bivar, D)
bivar_adjust
```

## Example - Monotone Regression

We demonstrate a GBL analysis with the <tt>cdiac</tt> data that ships with <tt>CVXR</tt>. The data contains Global Monthly and Annual Temperature Anomalies (degrees C, relative to the 1961-1990 mean) for years 1850-2015. This is plotted below.

```{r, echo = FALSE, warning = FALSE, message = FALSE}
data(cdiac)

ggplot(cdiac) +
  geom_point(aes(x = year, y = annual)) +
  xlab("Years") +
  ylab("Annual Temp Anomaly")
```

A traditional Bayes linear analysis would specify some prior belief structure and adjust by the data. For example

```{r}
n <- nrow(cdiac)
x <- cdiac$year

E_X <- E_D <- matrix(rep(0,n))

var_X <- cov_XD <- 0.3^2 * exp(-(as.matrix(dist(x))/25)^2) + 1e-6 * diag(rep(1, n))
var_D <- var_X + 0.1^2 * diag(rep(1, n))

bivar <- bs(E_X, E_D, cov_XD, var_X, var_D)

D <- cdiac$annual
temp_adjust <- adjust(bivar, D)
```

The result of this adjustment is shown below.

```{r, echo = FALSE}
plot_tbl <- tibble(
  year = as.numeric(x), 
  annual = as.numeric(temp_adjust$E_adj),
  sd = sqrt(diag(temp_adjust$var_adj))
)

ggplot(cdiac, aes(x = year, y = annual)) + 
  geom_point() +
  geom_line(data = plot_tbl) +
  geom_ribbon(
    data = plot_tbl, 
    aes(ymin = annual - 2*sd, ymax = annual + 2*sd),
    alpha = 0.2
  ) +
  xlab("Years") +
  ylab("Annual Temp Anomaly")

```

We can enforce the prior belief that this data is monotone increasing by specifying a <tt>gbl_bs</tt> object and adjusting by the data.

```{r}
gbl_bivar <- gbl_bs(
  E_X, E_D, cov_XD, var_X, var_D, 
  quote(list(CVXR::diff(Ec_adj) >= 0))
)

temp_const_adjust <- adjust(gbl_bivar, D)

```

The GBL inference is shown in blue.

```{r, echo = FALSE}
plot_gbl_tbl <- tibble(
  year = as.numeric(x), 
  annual = as.numeric(temp_const_adjust$Ec_adj),
  sd = sqrt(diag(temp_const_adjust$varc_adj))
)

ggplot(cdiac, aes(x = year, y = annual)) + 
  geom_point() +
  geom_line(data = plot_tbl) +
  geom_line(data = plot_gbl_tbl, color = "blue") +
  geom_ribbon(
    data = plot_tbl, 
    aes(ymin = annual - 2*sd, ymax = annual + 2*sd),
    alpha = 0.2
  ) + 
  geom_ribbon(
    data = plot_gbl_tbl, 
    aes(ymin = annual - 2*sd, ymax = annual + 2*sd),
    alpha = 0.2, fill = "blue"
  ) +
  xlab("Years") +
  ylab("Annual Temp Anomaly")

```

### The boring stuff

```{r}
sessionInfo()

```











