---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# bayeslinear

<!-- badges: start -->
<!-- badges: end -->

<tt>bayeslinear</tt> provides tools to perform a Bayes Linear analysis.

## Installation

<!-- You can install the released version of bayeslinear from [CRAN](https://CRAN.R-project.org) with:
``` r
install.packages("bayeslinear")
``` -->

<tt>bayeslinear</tt> is currently in development; install the latest version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("astfalckl/bayeslinear")
library(bayeslinear)
```
## Creating Belief Structures

A belief structure is the fundamental unit of information that we require to conduct a BL analysis blah blah blah. It requires the specification of $\mathbb{E}(X)$, $\mathbb{E}(D)$, $\mathrm{var}(X)$, $\mathrm{var}(D)$ and $\mathrm{cov}(X,D)$ to fully specify the inner product space. The methods in <tt>bayeslinear</tt> are based on the creation of a belief structure object through <tt>bs()</tt>. For example the one-dimensional example in Goldstein and Wooff (2007) is created as

```{r, echo = FALSE, warning = FALSE, message = FALSE}
setwd("/Users/lachlan/Documents/PostDoc/9_packages/bayeslinear")
library(devtools)
load_all()

```

```{r}
E_X <- 1
E_D <- 2

var_X <- var_D <- 1
cov_XD <- 0.6

one_dimension <- bs(E_X, E_D, cov_XD, var_X, var_D)
print(one_dimension)
```

## Adjusting Belief Structures

A <tt>bs</tt> object is adjusted by some data <tt>D</tt> via the <tt>adjust</tt> method. An adjusted belief structure <tt>adj_bs</tt> is returned. 

```{r}
D <- 0
adjust(one_dimension, D)
```





